---
title: "Microbiome data import & normalization"
author: "Jan Wälchli and Klaus Schlaeppi"
output:
  pdf_document:
    toc: yes
    toc_depth: 3
---

\pagebreak

```{r setup, include=FALSE, echo=F, message=F, warning=F}

#clear the objects from memory
rm(list=ls())

#knitr settings
knitr::opts_chunk$set(echo=TRUE, fig.align="center")

#set seed
set.seed(100)

#set source to file location
if (!require(rstudioapi)){install.packages("rstudioapi")}; library(rstudioapi)
setwd(dirname(getActiveDocumentContext()$path))

```

```{r paths, echo=F, message=F, warning=F}

#set import paths
paths <- list(DESIGN = c("../1_data/Rdesign.xlsx"),
              bASV_table = c("../../4_output/bacteria_ASV/bacteria_ASV.tab"),
              fASV_table = c("../../4_output/fungi_ASV/fungi_ASV.tab"),
              btaxa = c("../../4_output/bacteria_ASV/bacteria_taxa.tab"),
              ftaxa = c("../../4_output/fungi_ASV/fungi_taxa.tab"),
              interim = c("../1_data/interim")
              )

```

```{r libraries, echo=F, message=F, warning=F}

## load functions
source("functions/libraries.R")

## installs (if necessary) and loads libraries
libraries()

#load PCA data of soil chemistry
pca <- read.csv("../1_data/Exp_m18w19_PCA_soil_chem.csv")

#remove no longer used functions
rm(libraries)

```

```{r import design, echo=F, message=F, warning=F}

#import design file
DESIGN <- read_excel(paths$DESIGN)

#set rownames
rownames(DESIGN) <- DESIGN$Sample_ID

#Sample_type with colors
DESIGN$Sample_type <- factor(DESIGN$Sample_type, levels=c("root", "rhizosphere", "soil"))
DESIGN$cols_Sample_type <- as.character(DESIGN$Sample_type)
DESIGN$cols_Sample_type[DESIGN$Sample_type=="root"] <- "palegreen4"
DESIGN$cols_Sample_type[DESIGN$Sample_type=="rhizosphere"] <- "indianred3"
DESIGN$cols_Sample_type[DESIGN$Sample_type=="soil"] <- "chocolate4"

#collapsed color vector for each level
temp <- data.frame(DESIGN$Sample_type, DESIGN$cols_Sample_type)
temp <- plyr::ddply(temp, .variables="DESIGN.cols_Sample_type", .fun=unique)
level_cols_Sample_type <- as.character(temp[,2])
names(level_cols_Sample_type) <- temp[,1]

#Sample_type, Conditioning
DESIGN$Sample_type_Conditioning <- factor(paste(DESIGN$Sample_type, DESIGN$Conditioning, sep="_"), 
                                          levels=c("root_Bx+", "root_Bx-", "soil_Bx+", "soil_Bx-",
                                                   "rhizosphere_Bx+", "rhizosphere_Bx-"))

#Species
DESIGN$Species <- factor(DESIGN$Species, levels = c("Maize", "Wheat", "bare_soil"))
levels(DESIGN$Species) <- c("maize", "wheat", "bare-soil")

## Sample_type, Species
DESIGN$Sample_type_Species <- factor(paste(DESIGN$Sample_type, DESIGN$Species, sep="_"), 
                                          levels=c("root_maize", "root_wheat","rhizosphere_maize" , "rhizosphere_wheat", "soil_maize", "soil_wheat", "soil_bare-soil"))

DESIGN$cols_Sample_type_Species <- as.character(DESIGN$Sample_type_Species)
DESIGN$cols_Sample_type_Species[DESIGN$cols_Sample_type_Species=="root_maize"] <- "seagreen2"
DESIGN$cols_Sample_type_Species[DESIGN$cols_Sample_type_Species=="root_wheat"] <- "seagreen4"
DESIGN$cols_Sample_type_Species[DESIGN$cols_Sample_type_Species=="rhizosphere_maize"] <- "deepskyblue"
DESIGN$cols_Sample_type_Species[DESIGN$cols_Sample_type_Species=="rhizosphere_wheat"] <- "deepskyblue4"
DESIGN$cols_Sample_type_Species[DESIGN$cols_Sample_type_Species=="soil_maize"] <- "tan1"
DESIGN$cols_Sample_type_Species[DESIGN$cols_Sample_type_Species=="soil_wheat"] <- "tan3"
DESIGN$cols_Sample_type_Species[DESIGN$cols_Sample_type_Species=="soil_bare-soil"] <- "tan4"

## collapsed color vector for each level
temp <- data.frame(DESIGN$Sample_type_Species, DESIGN$cols_Sample_type_Species)
temp <- plyr::ddply(temp, .variables="DESIGN.cols_Sample_type_Species", .fun=unique)
level_cols_Sample_type_Species <- as.character(temp[,2])
names(level_cols_Sample_type_Species) <- temp[,1]

#Genotype
DESIGN$Genotype <- factor(DESIGN$Genotype, levels = c("W22", "bx1", "CH-Claro", "bare_soil"))
levels(DESIGN$Genotype) <- c("W22", "bx1", "CH-claro", "bare-soil")

#Conditioning with colors
DESIGN$Conditioning <- factor(DESIGN$Conditioning, levels = c("Bx+", "Bx-"))
DESIGN$cols_Conditioning <- as.character(DESIGN$Conditioning)
DESIGN$cols_Conditioning[DESIGN$Conditioning=="Bx+"] <- "gold2"
DESIGN$cols_Conditioning[DESIGN$Conditioning=="Bx-"] <- "darkgreen"

#collapsed color vector for each level
temp <- data.frame(DESIGN$Conditioning, DESIGN$cols_Conditioning)
temp <- plyr::ddply(temp, .variables="DESIGN.cols_Conditioning", .fun=unique)
level_cols_Conditioning <- as.character(temp[,2])
names(level_cols_Conditioning) <- temp[,1]

# Sample_type_Species_Conditioning
DESIGN$Sample_type_Species_Conditioning <- factor(paste(DESIGN$Sample_type, DESIGN$Species, DESIGN$Conditioning, sep="_"))

#run with colors
DESIGN$run <- as.factor(DESIGN$run)
DESIGN$cols_run <- as.character(DESIGN$run)
DESIGN$cols_run[DESIGN$run=="BE03"] <- "wheat1"
DESIGN$cols_run[DESIGN$run=="BE04"] <- "wheat3"

## collapsed color vector for each level
temp <- data.frame(DESIGN$run, DESIGN$cols_run)
temp <- plyr::ddply(temp, .variables="DESIGN.cols_run", .fun=unique)
level_cols_run <- as.character(temp[,2])
names(level_cols_run) <- temp[,1]

#add chem PCA 1 to design
pca$plot_id <- gsub("bx", "Bx", pca$plot_id)
rownames(pca) <- pca$plot_id 
pca_chem <- pca[,c("PCA_soil_chem_1", "PCA_soil_chem_4","plot_id")]
pca_for_design <- do.call("rbind", replicate(7, pca_chem, simplify = FALSE))
DESIGN$Sample_name <- as.factor(DESIGN$Sample_name)
DESIGN <- merge(pca_chem, DESIGN, by.y="Sample_name", by.x="plot_id")

#remove no longer needed files
rm(pca, pca_chem, pca_for_design)

```

```{r DAT import, echo=F, message=F, warning=F, include=F}

#Bacteria

#import ASV table
all_bDAT <- read.delim(paths$bASV, header=T, row.names=1, sep="\t")
all_bDAT <- t(all_bDAT)

#rename ASVs to bASVs for bacterial-ASVs
rownames(all_bDAT) <- gsub("ASV", "bASV", rownames(all_bDAT))

#adapt sample names to design file
colnames(all_bDAT) <- gsub("-","_", colnames(all_bDAT))
colnames(all_bDAT) <- gsub("_r1_F_filt.fastq","", colnames(all_bDAT))
colnames(all_bDAT) <- gsub("run","", colnames(all_bDAT))

#add primer names to DESIGN
DESIGN$bprimer <- ifelse(!is.na(DESIGN$BC_779F), paste(DESIGN$run_bac, DESIGN$BC_779F, DESIGN$run_bac, DESIGN$BC_1195R, sep="_"), NA)

#create bacterial-DESIGN
bDESIGN <- DESIGN [DESIGN$bprimer %in% colnames(all_bDAT), ]#keep only samples from bDAT
bDAT <- all_bDAT[, bDESIGN$bprimer]#order bDAT as bDESIGN


#Fungi

#import ASV table
all_fDAT <- read.delim(paths$fASV, header=T, row.names=1, sep="\t")
all_fDAT <- t(all_fDAT)

#rename ASV to bASVs for fungi OTUs
rownames(all_fDAT) <- gsub("ASV", "fASV", rownames(all_fDAT))

#adapt sample names to design file
colnames(all_fDAT) <- gsub("-","_", colnames(all_fDAT))
colnames(all_fDAT) <- gsub("_r1_F_filt.fastq","", colnames(all_fDAT))
colnames(all_fDAT) <- gsub("run","", colnames(all_fDAT))

#add primer names to DESIGN
DESIGN$fprimer <- ifelse(!is.na(DESIGN$BC_ITS1F), paste(DESIGN$run_fun, DESIGN$BC_ITS1F, DESIGN$run_fun, DESIGN$BC_ITS2, sep="_"), NA)

#create fungal-DESIGN
fDESIGN <- DESIGN [DESIGN$fprimer %in% colnames(all_fDAT), ]#keep only samples from fDAT
fDAT <- all_fDAT[, fDESIGN$fprimer]#order fDAT as fDESIGN

#remove no longer needed files
rm(all_bDAT, all_fDAT, DESIGN)

```


```{r bTAX import, echo=F, message=F, warning=F, include=F}

#Bacteria

#import taxonomy table
bTAX <- read.table(paths$btaxa, row.names=1, sep="\t", blank.lines.skip = FALSE)

#change cols to data type character
for (i in 1:6) {bTAX[,i] <- as.character(bTAX[,i])}

#rename bTAX
colnames(bTAX) <- c("kingdom", "phylum", "class", "order", "family", "genus")

#rename ASVs to bASVs for bacterial-ASVs
rownames(bTAX) <- gsub(">ASV","bASV", rownames(bTAX))

#remove non bacterial ASVs
r1 <- -which(bTAX$kingdom=="Eukaryota")
r2 <- -which(bTAX$phylum=="Cyanobacteria")
r3 <- -which(bTAX$family=="Mitochondria")
r4 <- -which(is.na(bTAX)) #rows all na
ASVs_to_remove <- c(r1,r2,r3,r4)
if(length(ASVs_to_remove)>0){bTAX <- bTAX[ASVs_to_remove ,]}
rm(r1, r2, r3, r4, ASVs_to_remove)

#rename unassigned ASVs
bTAX[is.na(bTAX)] <- "unassigned"

#add ASV_ID to taxonomy file
bTAX$ASV_ID <- rownames(bTAX)

#levels
bTAX$phylum <- as.factor(bTAX$phylum)
levels(bTAX$phylum)[levels(bTAX$phylum) < 2 ] <- "unassigned"

#define ASV colors by phylum (using the taxonomy file)
bTAX$labels <- as.character(bTAX$phylum)

#create separate taxonomy label specifying classes of Proteobacteria
bTAX$class <- as.factor(bTAX$class)
try(bTAX[ bTAX$class=="Alphaproteobacteria", ]$labels <- "Alphaproteobacteria")
try(bTAX[ bTAX$class=="Betaproteobacteria", ]$labels <- "Betaproteobacteria")
try(bTAX[ bTAX$class=="Gammaproteobacteria", ]$labels <- "Gammaproteobacteria")
try(bTAX[ bTAX$class=="Deltaproteobacteria", ]$labels <- "Deltaproteobacteria")

bTAX$labels <- as.factor(bTAX$labels)

```

```{r fTAX import, warning=F, echo=F,  message=F, include=F}

#Fungi

#import taxonomy table
fTAX <- read.table(paths$ftaxa, row.names=1, sep="\t", blank.lines.skip = FALSE)

#rename fTAX
colnames(fTAX) <- c("kingdom", "phylum", "class", "order", "family", "genus", "species")
fTAX$kingdom <- gsub("k__","", fTAX$kingdom )
fTAX$phylum <- gsub("p__","", fTAX$phylum )
fTAX$class <- gsub("c__","", fTAX$class )
fTAX$order <- gsub("o__","", fTAX$order )
fTAX$family <- gsub("f__","", fTAX$family )
fTAX$genus <- gsub("g__","", fTAX$genus )
fTAX$species <- gsub("s__","", fTAX$species )

#rename ASVs to fASVs for fungi ASVs
rownames(fTAX) <- gsub(">ASV","fASV", rownames(fTAX))

#remove non fungal ASVs
r1 <- -which(fTAX$kingdom=="Protista")
r2 <- -which(fTAX$kingdom=="Plantae")
r3 <- -which(fTAX$kingdom=="Protozoa")
r4 <- -which(fTAX$kingdom=="Animalia")
r5 <- -which(is.na(fTAX)) #rows all na
ASVs_to_remove <- c(r1,r2,r3,r4, r5)
if(length(ASVs_to_remove)>0){fTAX <- fTAX[ASVs_to_remove ,]}
rm(r1, r2, r3, r4, r5, ASVs_to_remove)

#rename unassigned ASVs
fTAX[is.na(fTAX)] <- "unassigned"

#add ASV_ID to taxonomy file
fTAX$ASV_ID <- rownames(fTAX)

#levels
fTAX$phylum <- as.factor(fTAX$phylum)
levels(fTAX$phylum)[levels(fTAX$phylum) < 2 ] <- "unassigned"

#define fASV colors by phylum (using the taxonomy file)
fTAX$labels <- as.character(fTAX$phylum)
fTAX$labels <- as.factor(fTAX$labels)
levels(fTAX$labels)

```

\pagebreak

# Data description

## Sequencing depth

```{r rarefaction threshold, echo=F, message=F, warning=F}

#remove very shallow sequenced samples
#In case we'd like to rarefy we define a rarefaction by the lowest seq depth

#Bacteria
b_rare <- sort(colSums(bDAT), decr=F)[5] #decide for threshold
b_rare <- reshape::round_any(b_rare, accuracy = 10, f=floor) #floor to 10

#remove samples with very low seq numbers
bsamples_rm <- names(which(colSums(bDAT) < b_rare))
bDESIGN <- bDESIGN[!(bDESIGN$bprimer %in% bsamples_rm),]
bDAT <- bDAT[,!(colnames(bDAT) %in% bsamples_rm)]
ASVs_rm <- names(which(rowSums(bDAT) == 0)) #rm ASVs with totSum=0
bDAT <- bDAT[!(rownames(bDAT) %in% ASVs_rm),]
bTAX <- bTAX[!(rownames(bTAX) %in% ASVs_rm),]

#Fungi
f_rare <- sort(colSums(fDAT), decr=F)[5] #decide for threshold
f_rare <- reshape::round_any(f_rare, accuracy = 10, f=floor) #floor to 10

#remove samples with very low seq numbers
fsamples_rm <- names(which(colSums(fDAT) < f_rare))
fDESIGN <- fDESIGN[!(fDESIGN$fprimer %in% fsamples_rm),]
fDAT <- fDAT[,!(colnames(fDAT) %in% fsamples_rm)]
ASVs_rm <- names(which(rowSums(fDAT) == 0)) #rm ASVs with totSum=0
fDAT <- fDAT[!(rownames(fDAT) %in% ASVs_rm),]
fTAX <- fTAX[!(rownames(fTAX) %in% ASVs_rm),]

```

We removed `r length(bsamples_rm)` bacterial samples with less than `r b_rare` sequences and `r length(fsamples_rm)` fungal samples with less than `r f_rare` sequences. We show the sum, range, median and total number of ASVs over all remaining samples. 

```{r seq numbers bacteria, echo=F, message=F, warning=F}

#create df to save sequences numbers
nseqs <- data.frame(taxa=c("Bacteria","Fungi"))

#Bacteria
nseqs$sum[1] <- sum(colSums(bDAT))
nseqs$min[1] <- min(colSums(bDAT))
nseqs$max[1] <- max(colSums(bDAT))
nseqs$median[1] <- median(colSums(bDAT))
nseqs$ASVs[1] <- nrow(bDAT)

#Fungi
nseqs$sum[2] <- sum(colSums(fDAT))
nseqs$min[2] <- min(colSums(fDAT))
nseqs$max[2] <- max(colSums(fDAT))
nseqs$median[2] <- median(colSums(fDAT))
nseqs$ASVs[2] <- nrow(fDAT)

#out
pander(nseqs)

#remove no longer used files
rm(nseqs)

```

### Figure 1 | Sequencing depth

```{r Boxplot sequences, fig.height=5, fig.width=11, echo=F, warning=F, message=F}

## Bacteria
df_bac <- data.frame(col_sum=colSums(bDAT, na.rm=T), 
                     colnames=colnames(bDAT), 
                     Genotype=bDESIGN$Sample_type_Species, 
                     #Compartment=DESIGN$Sample_type,
                     Conditioning=bDESIGN$Conditioning)
ylim2=boxplot.stats(df_bac$col_sum)$stats[c(1, 5)]

seq_nr_bac <- ggplot(data=df_bac, aes(x=Conditioning, y=col_sum, fill=Genotype)) + 
  geom_boxplot(position=position_dodge2(width=0.75, preserve="single") ) + 
  theme_bw() +
  theme(legend.position="none") +
  ylab("seq") +
  scale_fill_manual(values=level_cols_Sample_type_Species) +
  coord_cartesian(ylim=ylim2*1.05) +  
  ggtitle("Bacteria: Sequencing depth")

## Fungi
df_fun <- data.frame(col_sum=colSums(fDAT, na.rm=T), 
                     colnames=colnames(fDAT), 
                     Genotype=fDESIGN$Sample_type_Species, 
                     #Compartment=DESIGN$Sample_type,
                     Conditioning=fDESIGN$Conditioning)

ylim2=boxplot.stats(df_fun$col_sum)$stats[c(1, 5)]

seq_nr_fun <- ggplot(data=df_fun, aes(x=Conditioning, y=col_sum, fill=Genotype)) + 
  geom_boxplot(position=position_dodge2(width=0.75, preserve="single") ) + 
  theme_bw() +
  theme(legend.position="none") +
  ylab("seq") +
  scale_fill_manual(values=level_cols_Sample_type_Species) +
  coord_cartesian(ylim=ylim2*1.05) +  
  ggtitle("Fungi: Sequencing depth")

```

```{r Barplot sequences, fig.height=5, fig.width=11, echo=F, warning=F, message=F}

## Bacteria
df_bac$colSum <- colSums(bDAT)
df_bac <- df_bac[order(df_bac$colSum, decreasing = T),]
df_bac$x <- 1:nrow(df_bac)

samples_bac <-  ggplot(df_bac, aes(x=x, y=colSum, fill=Genotype, color=Genotype))+
                geom_bar(aes(width=1), size=0.1, stat = "identity", col="black")+
                #scale_y_continuous(trans = 'log2')+
                scale_fill_manual(values=level_cols_Sample_type_Species) +
                scale_colour_manual(values = level_cols_Sample_type_Species) +
                xlab("Samples")+
                ylab("seq / sample")+
                theme_bw() +
                theme(axis.text.x=element_blank(), axis.ticks.x=element_blank(),
                      strip.background = element_blank(), strip.text.x = element_blank()) #, legend.position = "none"

## Fungi
df_fun$colSum <- colSums(fDAT)
df_fun <- df_fun[order(df_fun$colSum, decreasing = T),]
df_fun$x <- 1:nrow(df_fun)

samples_fun <-  ggplot(df_fun, aes(x=x, y=colSum, fill=Genotype, color=Genotype))+
                geom_bar(aes(width=1), size=0.1, stat = "identity", col="black")+
                #scale_y_continuous(trans = 'log2')+
                scale_fill_manual(values=level_cols_Sample_type_Species) +
                scale_colour_manual(values=level_cols_Sample_type_Species) +
                xlab("Samples")+
                ylab("seq / sample")+
                theme_bw() +
                theme(axis.text.x=element_blank(), axis.ticks.x=element_blank(),
                      strip.background = element_blank(), strip.text.x = element_blank()) #, legend.position = "none"

```

```{r Figure 1, fig.height=8, fig.width=11, echo=F, warning=F, message=F}

# library(cowplot)

legend <- get_legend(samples_bac + theme(legend.position="bottom") + labs(fill="Genotype"))

p <- plot_grid(seq_nr_bac + theme(legend.position="none"),
                   seq_nr_fun + theme(legend.position="none"),
                  samples_bac + theme(legend.position="none") + labs(colour=""),
                  samples_fun + theme(legend.position="none") + labs(colour=""),
                  labels=c("A","B"),
                  align = "hv",
                  rel_heights=c(1, .5),
                  nrow=2)

p <- plot_grid(p, legend, nrow=2, rel_heights=c(1, .1))
print(p)

```

\pagebreak

## Rarefaction curve

We checked if the sequence depth is enough to capture the microbial diversity by plotting the rarefaction curve for each sample.

### Figure 2 | Rarefaction curve

```{r rarefaction, eval=T, echo=F, message=F, warning=F, fig.height=7, fig.width=14}

#Bacteria

#rarefy
pdf("/dev/null") #prevents drawing of standard rarefaction plot (we will create our own with colors)
brarefaction <- vegan::rarecurve(t(bDAT), step=50, 7000, label = FALSE)
names(brarefaction) <- bDESIGN$Sample_ID
invisible(dev.off())

#long-transformation.
protox <- mapply(FUN = function(x, y) {
  mydf <- as.data.frame(x)
  colnames(mydf) <- "value"
  mydf$species <- y
  mydf$subsample <- attr(x, "Subsample")
  mydf
}, x = brarefaction, y = as.list(names(brarefaction)), SIMPLIFY = FALSE)
brarefaction_long <- do.call(rbind, protox)
rownames(brarefaction_long) <- NULL  # pretty

#add Sample_type_Species
brarefaction_long$Sample_type_Species <- NA
for (i in 1:nrow(brarefaction_long)) {
  brarefaction_long$Sample_type_Species[i] <- as.character(bDESIGN$Sample_type_Species[bDESIGN$Sample_ID == brarefaction_long$species[i]])
}

#add Conditioning
brarefaction_long$Conditioning <- NA
for (i in 1:nrow(brarefaction_long)) {
  brarefaction_long$Conditioning[i] <- as.character(bDESIGN$Conditioning[bDESIGN$Sample_ID == brarefaction_long$species[i]])
}

#plot
brareplot <- ggplot(brarefaction_long, aes(x = subsample, y = value, group = species)) +
                  theme_bw() +
                  geom_line(aes(color=Sample_type_Species, lty=Conditioning))+
                  scale_color_manual(values = level_cols_Sample_type_Species)+
                  xlab("Sequencing Depth")+
                  ylab("Number of ASVs")+
                  ggtitle("Bacteria")


#Fungi

#rarefy
pdf("/dev/null") #prevents drawing of standard rarefaction plot (we will create our own with colors)
frarefaction <- vegan::rarecurve(t(fDAT), step=50, 7000, label = FALSE)
names(frarefaction) <- fDESIGN$Sample_ID
invisible(dev.off())

# long-transformation.
protox <- mapply(FUN = function(x, y) {
  mydf <- as.data.frame(x)
  colnames(mydf) <- "value"
  mydf$species <- y
  mydf$subsample <- attr(x, "Subsample")
  mydf
}, x = frarefaction, y = as.list(names(frarefaction)), SIMPLIFY = FALSE)
frarefaction_long <- do.call(rbind, protox)
rownames(frarefaction_long) <- NULL  # pretty

#add Sample_type_Species
frarefaction_long$Sample_type_Species <- NA
for (i in 1:nrow(frarefaction_long)) {
  frarefaction_long$Sample_type_Species[i] <- as.character(fDESIGN$Sample_type_Species[fDESIGN$Sample_ID == frarefaction_long$species[i]])
}

#add Conditioning
frarefaction_long$Conditioning <- NA
for (i in 1:nrow(frarefaction_long)) {
  frarefaction_long$Conditioning[i] <- as.character(fDESIGN$Conditioning[fDESIGN$Sample_ID == frarefaction_long$species[i]])
}

#plot
frareplot <- ggplot(frarefaction_long, aes(x = subsample, y = value, group = species)) +
                  theme_bw() +
                  geom_line(aes(color=Sample_type_Species, lty=Conditioning))+
                  scale_color_manual(values = level_cols_Sample_type_Species)+
                  xlab("Sequencing Depth")+
                  ylab("Number of ASVs")+
                  ggtitle("Fungi")


#combine plots (package:cowplot)
legend <- get_legend(brareplot + theme(legend.position="bottom") + labs(col="Sampletype-Species"))
p <- plot_grid(brareplot + theme(legend.position="none"),
               frareplot + theme(legend.position="none"),
                  labels=c(" "," "), align = "hv", nrow=1)

p <- plot_grid(p, legend, nrow=2, rel_heights=c(1, .1))
print(p)

```

\pagebreak

# Data normalization

To decide on how to normalize the data we followed the recommendation of Weiss et al. (2017, Microbiome Journal) and we inspected whether there are differences in sequencing depths between the different sample groups utilizing the non-parametric Kruskal-Wallis Test. 

## Asymptotic Kruskal-Wallis test

We tested different sequencing depths for bacteria and fungi.

```{r seq numbers per group fungi, echo=F, message=F, warning=F}

## stats bac
DAT_seq_krusk <- kruskal_test(colSums(bDAT) ~ bDESIGN$Sample_type_Species_Conditioning)  # library(coin)
DAT_seq_krusk

## stats fun
DAT_seq_krusk <- kruskal_test(colSums(fDAT) ~ fDESIGN$Sample_type_Species_Conditioning)  # library(coin)
DAT_seq_krusk

```

*Conclusion:* We found significant differences in sequencing depths between the different sample groups for bacteria and fungi. Therefore we rarefied the data for bacteria and fungi to equalize sequence differences! We defined the rarefaction threshold for bacteria to `r b_rare` sequences per sample and for fungi to `r f_rare` sequences.

```{r DAT rare, warning=F, echo=F}

#Bacteria
#rarefaction (package vegan)
bDAT_rare <- t(rrarefy(t(bDAT), b_rare))
bDAT_rare <- bDAT_rare[rowSums(bDAT_rare) > 0,]  # removal of rows with 0 values

#Fungi
#rarefaction (package vegan)
fDAT_rare <- t(rrarefy(t(fDAT), f_rare))
fDAT_rare <- fDAT_rare[rowSums(fDAT_rare) > 0,]  # removal of rows with 0 values

```

\pagebreak

# Final number of samples 

We ended up with the following number of samples per treatment for the analysis.  

```{r number of samples, warning=F, message=F, echo=F}

#nsamples
btable <- table(bDESIGN$Sample_type_Species_Conditioning)
ftable <- table(fDESIGN$Sample_type_Species_Conditioning)

#Concatenate
sample_number <- data.frame(bac=btable, fun=ftable)
if(all(sample_number$bac.Var1 == sample_number$fun.Var1)){
  sample_number <- sample_number[c("bac.Var1","bac.Freq","fun.Freq")]
  names(sample_number) <- c("Sample", "Bacteria", "Fungi")
} else{"Check order"}

#out
pander(sample_number, caption = "Sample profile")


```

```{r export RDA files, echo=F, warning=F, message=F}

#create directory
dir.create(paths$interim)

## set output directory
setwd(paths$interim)

#save objects needed in the following scripts as RDS
saveRDS(bDESIGN, "bDESIGN.RDS")
saveRDS(fDESIGN, "fDESIGN.RDS")

saveRDS(bDAT, "bDAT.RDS")
saveRDS(fDAT, "fDAT.RDS")
saveRDS(b_rare, "b_rare.RDS")
saveRDS(f_rare, "f_rare.RDS")
saveRDS(bDAT_rare, "bDAT_rare.RDS")
saveRDS(fDAT_rare, "fDAT_rare.RDS")

saveRDS(bTAX, "bTAX.RDS")
saveRDS(fTAX, "fTAX.RDS")

saveRDS(level_cols_Sample_type, "level_cols_Sample_type.RDS")
saveRDS(level_cols_Sample_type_Species, "level_cols_Sample_type_Species.RDS")
saveRDS(level_cols_Conditioning, "level_cols_Conditioning.RDS")

```

\pagebreak

